---
title: "Full Analysis Pipeline"
author: "Christian Pascual"
date: "11/1/2021"
output: pdf_document
---

# Preparation For Building Networks

```{r}
# Vector to stop future from arcing into past
past = c("p_steps", "p_diff", "p_gaz")
inputs = c("steps", "diff")
prev_inputs = c("p_steps", "p_diff")
ema_preds = c("busypred", "stresspred", "typicalpred")
ema_actual = c("hecticactual", "stressactual", "typicalactual")

# Blacklist for preventing certain arcs from  being created in networks
bl = matrix(c(
  
  # The future shouldn't affect the past
  c(rbind("steps", past)),
  c(rbind("diff", past)),
  c(rbind("gaz", past)),
  
  "p_steps", "p_diff",
  "p_gaz", "p_diff",
  
  # Preventing post-steps from arcing into pre-steps
  c(rbind("gaz", inputs)),
  c(rbind("p_gaz", prev_inputs)),
  
  # goal difficulty should not affected by anything
  c(rbind(inputs, "diff")),
  c(rbind(prev_inputs, "diff")),
  c(rbind(ema_preds, "diff")),
  c(rbind(ema_actual, "diff")),
  
  # Predictions should not come after the days events
  c(rbind("hecticactual", ema_preds)),
  c(rbind("stressactual", ema_preds)),
  c(rbind("typicalactual", ema_preds)),
  c(rbind("steps", ema_preds)),
  c(rbind("diff", ema_preds)),
  c(rbind("gaz", ema_preds)),
  c(rbind("hecticactual", past)),
  c(rbind("stressactual", past)),
  c(rbind("typicalactual", past)),
  c(rbind("steps", "sleep")),
  c(rbind("diff", "sleep")),
  c(rbind("gaz", "sleep")),
  c(rbind("busypred", past)),
  c(rbind("stresspred", past)),
  c(rbind("typicalpred", past))
  
),
ncol = 2, byrow = TRUE, 
dimnames = list(NULL, c("from", "to")))
```

# Load In Raw Just Walk Data

```{r, message = F }
raw_just_walk = here::here("data", "Data_JustWalk.csv") %>% 
  read_csv() %>% 
  select(participant:self_e_tminus1) %>% 
  mutate(
    participant = as.character(participant),
  ) %>% 
  group_by(participant) %>% 
  mutate(
    prev_goal_difficulty = lag(goal_difficulty, 1),
    prev_actual_steps = lag(actual_steps, 1)
  ) %>% 
  ungroup %>% 
  filter(!is.na(goal_difficulty_tminus1))
  
# Prepare the data for use in the networks
just_walk = raw_just_walk %>% 
  select(
    participant,
    day,
    steps = actual_steps,
    diff = goal_difficulty,
    gaz = g_achiev_z, 
    p_steps = prev_actual_steps,
    p_diff = prev_goal_difficulty,
    p_gaz = g_achiev_z_tminus1
  )
  
# Load in the imputed data and match
imputed_ema = read_csv("../data/imputed_ema.csv") %>% 
  mutate( participant = as.character(participant) )
  
full_just_walk = just_walk %>% 
  left_join(imputed_ema, by = c("participant", "day")) %>% 
  drop_na()
```

# Create Networks Using First N Days Of Data

```{r}
R = 1000 # number bootstrap networks too average from
limited_individual_networks = full_just_walk %>% 
  group_by(participant) %>% 
  summarize( 
    start = min(day),
    last = max(day) 
  ) %>% 
  mutate(
    n_days = map2(start, last, function(s, t) { tibble(n_days = (s + 1):t) })
  ) %>% 
  unnest_wider(n_days) %>% 
  unnest(n_days) %>% 
  select(-start, -last) %>% 
  mutate(
    # Filter the data down to a limited set of days
    lim_data = map2(participant, n_days, function(p, n) {
      full_just_walk %>% 
        filter(participant == p, day <= n) %>% 
        select(-participant, -day)
    }),
    # Create the network based off the limited data
    lim_network = pmap(list(participant, lim_data, n_days), function(p, d, n) {
      
      # Progress bar for sanity:
      print(paste0("ID: ", p, ", Network for ", n, " days"))
      
      arcs = boot.strength(d, 
                           algorithm = "hc",
                           algorithm.args = list("blacklist" = bl),
                           R = R)
      
      mean_arcs = mean(arcs)
      averaged.network(mean_arcs)
      
    })
  ) 

limited_aggregate_networks = tibble(
    participant = "Aggregated",
    n_days = 17:93, # range of days in the data
    lim_data = map(n_days, function(n) {
      
      full_just_walk %>% 
        filter(day <= n) %>% 
        select(-participant, -day)
      
    }),
    lim_network = map(lim_data, function(d) {
      
      arcs = boot.strength(d, 
                           algorithm = "hc",
                           algorithm.args = list("blacklist" = bl),
                           R = R)
      
      mean_arcs = mean(arcs)
      averaged.network(mean_arcs)
      
    })
  )

limited_networks = bind_rows(limited_individual_networks,
                             limited_aggregate_networks)
save(limited_networks, file = "data/limited_networks_final.rda")
```

```{r}
# check: what do the networks look like
NetworkDisplayApp()
```

Note To Self:

- some arcs were created in the averaged network that would produce cycles
   - these were presumed to be ignored + not included in the plot

# Calculate Change in BIC For Each Arc In Each Network

- averaged.network() sometimes returns undirected edges, so some processing is needed to remove all of them in a principled way

```{r}
bic_analysis = limited_networks %>% 
  mutate(
    final_lim_network = map2(lim_network, lim_data, function(ntwk, d) {
      
      # Remove all undirected arcs (happens with small data)
      undirected = undirected.arcs(ntwk)
      if (nrow(undirected) > 0) {
        # Undirected arcs come in pairs, so opt to remove even indices
        for (i in seq(1, nrow(undirected)/2, by = 1)) {
          
          ntwk = drop.arc(ntwk, from = undirected[i,1], to = undirected[i,2])
        }
      }
      
      ntwk
      
    }),
    bic_change = map2(final_lim_network, lim_data, function(ntwk, d) { 
      
      arcstrength = arc.strength(ntwk, d) 
      
      full_bic = score(ntwk, d)
      arcs = arcs(ntwk)
      n_arcs = arcs(ntwk) %>% nrow()
      
      # If no arcs there should be no table
      if (n_arcs == 0) { 
        return(data.frame(from = character(),
                          to = character(),
                          strength = double(),
                          full_bic = double(), 
                          drop_bic = double(), 
                          change_bic = double()))
      }
      
      bics = double(n_arcs)
      change_bics = double(n_arcs)
      
      for (i in 1:n_arcs) {
        
        # Remove an arc from the averaged network
        small_ntwk = drop.arc(ntwk, from = arcs[i,][1], to = arcs[i,][2])
        
        # Calculate new BIC of the smaller network when arc is removed
        # The lower the better
        bics[i] = score(small_ntwk, d) 
        change_bics[i] = full_bic - score(small_ntwk, d)
        
      }
      
      cbind(arcstrength,
            full_bic = full_bic, 
            drop_bic = bics, 
            change_bic = change_bics)

      })
  ) 

bic_data = bic_analysis %>% 
  select(participant, n_days, bic_change) %>% 
  unnest_wider(bic_change) %>% 
  unnest(c("from", "to", "strength", "full_bic", "drop_bic", "change_bic")) %>% 
  mutate(
    arcstr = map2_chr(from, to, function(f, t) {
      paste0(f, "->", t)
    })
  )

save(bic_data, file = "data/bic_data_final.rda")
```

Note To Self:

- The unnesting removes all of the networks that have no arcs in them (ie early arcs)

# Calculate The True And False Positives With Time

```{r}
last_days = bic_analysis %>% 
  group_by(participant) %>% 
  summarize( max_days = max(n_days) )

full_networks = bic_analysis %>% 
  left_join(last_days, by = "participant") %>% 
  filter(n_days == max_days) %>% 
  select(participant, full_network = final_lim_network)

development_data = bic_analysis %>% 
  left_join(full_networks, by = "participant") %>% 
  select(participant, n_days, final_lim_network, full_network) %>% 
  mutate(
    comp =  map2(final_lim_network, full_network, function(lim, fin) {
      compare(target = fin, current = lim)
    }),
    comparcs =  map2(final_lim_network, full_network, function(lim, fin) {
      compare(target = fin, current = lim, arcs = T)
    }),
    n_arcs_full = map_dbl(full_network, function(ntwk) {nrow(arcs(ntwk))}),
    true_positive = map_dbl(comp, function(cc) { cc$tp }),
    false_positive = map_dbl(comp, function(cc) { cc$fp }),
    false_negative = map_dbl(comp, function(cc) { cc$fn }),
    true_positive_arcs = map(comparcs, function(cc) { as_tibble(cc$tp) }),
    false_positive_arcs = map(comparcs, function(cc) { as_tibble(cc$fp) }),
    false_negative_arcs = map(comparcs, function(cc) { as_tibble(cc$fn) }),
    prop_tp = true_positive / n_arcs_full
  )

save(development_data, file = "data/network_development_final.rda")
```

# Match Up BIC Changes To TP, FP, and FN

```{r}
check = development_data %>% 
  select(participant, n_days, false_positive_arcs) %>% 
  unnest_wider(false_positive_arcs) %>% unnest(c("from", "to")) %>% 
  mutate(
    arcstr = map2_chr(from, to, function(f, t) {
      paste0(f, "->", t)
    })
  ) %>% 
  left_join(bic_data %>% select(participant, n_days, arcstr, change_bic),
            by = c("participant", "n_days", "arcstr"))

c("true_positive_arcs", "false_positive_arcs", "false_negative_arcs")
```

# Create HeatMap Data

```{r}
# heat map for checking the presence of absence of an arc on a given day
# get all arcs that were ever made for all networks
all_arcstrings = bic_data %>% pull(arcstr) %>% unique
ids = bic_data %>% pull(participant) %>% unique
days = bic_data %>% pull(n_days) %>% unique

# All possible participant, arc and day number combinations
p = expand.grid(ids, all_arcstrings, days) %>% 
  as_tibble()
colnames(p) = c("participant", "arcstr", "n_days")
p = p %>% mutate(arcstr = unlist(arcstr))

# Tibble of
present = bic_data %>% 
  select(participant, n_days, arcstr, change_bic) %>% 
  mutate( 
    arcstr = unlist(arcstr)
  )

id_max_day = bic_data %>% 
  group_by(participant) %>% 
  summarize( last = max(n_days) )

heat = p %>% 
  left_join(present, by = c("participant", "n_days", "arcstr")) %>% 
  left_join(id_max_day, by = "participant") %>% 
  filter(n_days <= last) %>% 
  mutate(
    present = if_else(is.na(change_bic), 0, 1)
  )

save(heat, file = "data/heatmap_data_final.rda")
```

# checking how long it takes to learn a network

```{r}
# Time to learn a full network
a = development_data %>% 
  transmute(
    time = n_days,
    prop_present = true_positive / n_arcs_full,
    above_50 = (prop_present >= 0.50) %>% as.numeric,
    above_75 = (prop_present >= 0.75) %>% as.numeric,
    above_90 = (prop_present >= 0.90) %>% as.numeric
  ) %>% 
  group_by(time) %>% 
  summarize(
    `50% learned` = mean(above_50),
    `75% learned` = mean(above_75),
    `90% learned` = mean(above_90)
  ) %>% 
  pivot_longer(
    `50% learned`:`90% learned`,
    names_to = "prop",
    values_to = "val"
  )

a %>% 
  ggplot(aes(x = time, y = val, color = prop)) +
  geom_line() + 
  labs(
    x = "Days used to learn network",
    y = "Proportion of sample who have learned\n a portion of full network"
  )
```

```{r}

```

